What, if anything, is epsilon?

### 2014-03-01 epsilon/paper.tex
Progress on paper, tables, etc.

In mathematics, $\epsilon$ usually refers to the
$\epsilon--\delta$ formulation of limits. This is pretty simple and a
reminder appears in Figure~\ref{fig:epsilondelta}. This paper is not
about that kind of math.

@@

Most programmers find this subject too tedious and simply pick a
number that seems pretty small. Thus in practice, $\epsilon$ is an
application-specific choice. Of well-known ``constants,''
$\epsilon$ may be the least agreed-upon, with values seen in the wild
spanning {\em 50 orders of magnitude}. This paper explores the practical
values of $\epsilon$ in real software.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.90 \linewidth]{epsilondelta.pdf}
\end{center}\vspace{-0.1in}
\caption{explain the graph}
\label{fig:epsilondelta}
\end{figure}

@@

>>>>

In mathematics, $\epsilon$ usually refers to the
$\epsilon$--$\delta$ formulation of limits. This is pretty simple and a
reminder appears in Figure~\ref{fig:epsilondelta}. This paper is not
about that kind of math.

@@

Most programmers find this subject too tedious and simply pick a
number that seems pretty small. Thus in practice, $\epsilon$ is an
application-specific choice. Of well-known ``constants,''
$\epsilon$ may be the least agreed-upon, with values seen in the wild
spanning {\em 300 orders of magnitude}. This paper explores the practical
values of $\epsilon$ in real software.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.90 \linewidth]{epsilondelta.pdf}
\end{center}\vspace{-0.1in}
\caption{ The curvy line is the function $\frac{\sin x}{x}$, which is
  undefined at $0$. However, its limit as $x$ approaches $0$ is $1$.
  The $\epsilon$--$\delta$ formulation of the limit is this: For any
  positive choice of $\epsilon$, there exists some $\delta$ such that
  $\frac{\sin (0 + \delta)}{0 + \delta}$ is less than $1 + \epsilon$.
  In other words, for an arbirarily small error (your choice), I
  can produce a delta from $0$ (my choice) that brings the result
  within the error of the limit. This has nothing to do with the
  subject of the paper.}
\label{fig:epsilondelta}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99 \linewidth]{chart-c-double}
\end{center}\vspace{-0.1in}
\caption{ Values for {\tt const double epsilon} in the C programming
  language. In this chart, the blue (lower) bars are the distinct values
  of epsilon seen. The vertically aligned red (upper) bar is its count.
  Epsilon values are plotted on a logarithmic scale, where the minimum
  observed value $\log(-708)$ is $303 \times 10^{-308}$, and the largest
  $\log(1.609)$ is $5$. 
  %
  Notes: One programmer used the value \tt{-1e10}, which is
  -10,000,000,000, problably meaning \tt{1e-10}. This value was excluded
  because it has no real logarithm.
}
\label{fig:cdouble}
\end{figure}

\begin{figure}[ht]
\begin{center}
\begin{tabular}{l}
\verb+0.5/ELEC_REST_ENERGY+ \\
\verb+alpha/beta+ \\
\verb+4 / MULT32+ \\
\verb+exact_epsilon(true)+ \\
\verb+fmass_Epsilon * EPS_EXTRA+ \\
\verb+((Lj_Parameters*) parameters)->+ \\
\verb+scalar_traits<+ \\
\verb+EpsArray[prec]+ \\
\verb+hfwfn_->+ \\
\verb+fl.net_.opt_.epsilon+ \\
\verb+Tolerance+ \\
\end{tabular}
\end{center}\vspace{-0.1in}
\caption{ Other uninterpretable values of {\tt const double epsilon}
  in C. Who knows what these are supposed to be?}
\label{fig:cdoubleuninterpretable}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99 \linewidth]{chart-c-float}
\end{center}\vspace{-0.1in}
\caption{ Values for {\tt const float epsilon} in the C programming
  language. An $x$--$y$ scatter plot where the $y$ coordinate is the
  count of the number of times that specific value occurred, and the
  $x$ coordinate is the log of the value. Values take on a less
  extreme range than with type {\tt double}, naturally, ranging ``only''
  31 orders of magnitude from $3.9 \times 10^{-31}$ to $0$.}
\label{fig:cfloat}
\end{figure}

\begin{figure}[ht]
\begin{center}
% \includegraphics[width=0.99 \linewidth]{chart-c-double}
\end{center}\vspace{-0.1in}
\caption{
  Values for {\tt const double epsilon} and {\tt constexpr
    double epsilon} in the C++ programming language. The {\tt
    constexpr} qualifier, a new feature of C++11, is used very rarely
  (less than 1\% of the time).
  Notes:
  Five times, the computed value is actually equal to 0.0,

  The value {\tt pow(10,-13)} is annotated in German
  ``Genauigkeitsziel bei der Nullstellensuche,'' or ``Accuracy goal in
  the search for zeros.''
}
\label{fig:cppdouble}
\end{figure}
